---
  title: "01-assignment.Rmd"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Assignment 1

For your first assignment, I want you to create a file called 01-assignment_<lastname>.Rmd .
elements:
  1. A sentence that says “Hello, World”



2. R output that summarizes one of the variables in the colllege.Rdata
dataset



Submit it under assignment 1, using the format 01-assignment-<lastname>.Rmd. All assignments should be turned in using this format. Since my last name is Doyle, I would use 01-assignment-doyle.Rmd as my file name. Unless your name is also Doyle, you should use a different name.

Stretch Items

If you have extra time, you can do the following:
  
  1. Calculate the average earnings for individuals at the most selective colleges, then compare that with individuals at the least selective colleges in the dataset.

2. Find a way to determine whether colleges with very high SAT scores tend to be larger or smaller than colleges with low SAT scores.

3. Plot the relationship between cost and debt. What do you see? Does this surprise you?
  
  4. Now, provide separate plots for cost and debt by control of the institution.
