---
title: "Introduction to Data Science"
output:
  word_document: default
  html_document: default
---

Introduction
=============

We have entered a time in which vast amounts of data are more widely
available than ever before. At the same time, a new set of tools has
been developed to analyze this data and provide decision makers with
information to help them accomplish their goals. Those who engage with
data and interpret it for organizational leaders have taken to calling
themselves data scientists, and their craft data science. Other terms
that have come into vogue are "Big Data," "Predictive Analytics" and
"Data Mining." These can seem to be mysterious domains. The point of
this class is to demystify much of this endeavor for individuals who
will be organizational leaders.

The class is structured around developing students' skills in three
areas: getting data, analyzing data to make predictions, and
presenting the results of analysis. For each area, the subtopics are
as follows:

Getting Data Topics
--------------------
* Tools of the Trade: R and Rstudio
* Working with pre-processed data and flat files
* Getting data from the web: webscraping, using forms, using Application Programming Interfaces
* Using databases

Analyzing Data Topics
----------------------
* Descriptives and conditional means
* Regression
* Supervised learning: classification
* Unsupervised learning: K-means and nearest neighbors clustering
* Cross Validation

Presenting Data Analysis Topics
---------------------------------
* Descriptives: histograms, density plots, bar plots, dot
  plots
* Scatterplots
* Lattice graphics and small multiples
* Interactive graphics
* Communicating results effectively

Evaluation
===========
Students will be evaluated based in two areas: weekly Problem sets and
the final project. 

* Problem sets: 65% Each week I will assign a problem set for
  students to complete. These problem sets will be assigned on Monday,
  and will be due the next Sunday night at 11:59:59 pm. No late
  problem sets will be accepted. Each problem set will be graded on a
  100 point scale. Your lowest grade will be dropped. 
  
  There will be 13 assigned problem sets, with each problem set worth 100 points. The lowest grade will be dropped, meaning that you will be graded on 12 of these problem sets. The grading standards will be as follows:
  
  50=turned in problem set, did not attempt most of the problems
  
  75= turned in problem set, attempted most of the problems
  
  100= turned in problem set, attempted all of the problems
  
  Note that your grade on problem sets does not depend on you being correct on all problems, but simply making a serious attempt to answer all of the problems. 

* Final Project 35%: During the course of the semester you will
  work on a final assignment utilizing your skills as a data
  analyst.  
  
There will be four progress reports for the final project, each of which will be worth 12.5% of the final  grade for the project. No late progress reports will be accepted. The final product will account for the remaining 50%. No late final products will be accepted. 
  
There are two parts to the final project grade, your progress reports and the product itself. Each project report is worth 12.5% of the grade. These will be graded based on completion of the enumerated steps in the report, and will be graded on a 100 point scale.
  
The final product will be worth 50% of your final project grade. A rubric will be supplied containing the relevant grading information for this final product. 
  

Texts
======

*Required Texts*

We will have two texts for the course. The first is Hadley Wickham's book, 
[R for Data Science](http://r4ds.had.co.nz/). 

Wickham, H., & Grolemund, G. (2016). R for data science: import, tidy, transform, visualize, and model data. San Francisco: O'Reilly Media, Inc.

The other text is Nate Silver's *Signal and the Noise*. 

Silver,N. (2012). *The signal and the noise: Why so many predictions fail-but some don’t*. New York: Penguin. 

Software
==========
We will use only free, [open source](https://opensource.org/osd-annotated) software in this course. 

We will use  [R](https://cran.r-project.org/), an open-source data analytic
platform for all analysis. R appears to be the most widely used data analysis software in data science. We will utilize [Rstudio](https://www.rstudio.com/home/) as our integrated development environment (IDE) for R. 

Honor Code Statement
===============

All assignments for this class, including weekly problem sets and the final project, are to be conducted under the obligations set out in Vanderbilt's Honor Code. Please click [here](http://www.vanderbilt.edu/student_handbook/the-honor-system/)to review the honor code.

There will be two quite different standards for completing the
problem sets and the final project. 

*problem sets* You may collaborate with anyone and you may utilize
  any resource you wish to complete these problem sets. 
  
*Final Project* All of the work on the final assignment must be
  your own. Anyone's work that you reference should be cited, as
  usual. All data that you do not personally collect must be cited, as with any other resource. 

If you have any questions at all about the honor code or how it will
be applied, ask me right away. 

Schedule
=========
```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-1
```
## Week `r i`: Topic for the Week: Getting Data 1-- Tools of the Trade

*Resources*

Wickham: [Introduction](http://r4ds.had.co.nz/introduction.html), [Explore: Introduction](http://r4ds.had.co.nz/explore-intro.html), [Workflow: basics](http://r4ds.had.co.nz/workflow-basics.html), [Workflow: projects](http://r4ds.had.co.nz/workflow-projects.html)

Silver, Chapters 1-4

R Intro and Resources

[Download R](https://cran.r-project.org/)

[R Basics](https://www.datacamp.com/home)

[Download Rstudio](https://www.rstudio.com/products/rstudio/download3/)
You want the "Desktop" version, free license

[Rstudio Intro and Resources](https://www.rstudio.com/online-learning/)

*Lesson Notes*

Chapter 1, Introduction: `01-intro.Rmd`.

*Synchronous session* R Basics, "verbs" of data wrangling

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i`  Topic for the Week: Analyzing Data: Conditional Means

*Resources*
  
Wickham: [Data transformation](http://r4ds.had.co.nz/transform.html)  
Silver, Chapters 5-9, 12-13

*Lecture Notes*

Chapter 2, Condtional Means: `02-conditional_means.Rmd`.

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Synchronous Section: Conditional Means*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```


## Week `r i` Topic for the Week: Presenting Data-- Descriptives

*Resources*

Wickham: [Data visualization](http://r4ds.had.co.nz/data-visualisation.html)

[Data transformation](http://r4ds.had.co.nz/exploratory-data-analysis.html)

[Cookbook for R:Bar and Line Graphs](http://www.cookbook-r.com/Graphs/Bar_and_line_graphs_%28ggplot2%29/)

[Cookbook for R: Plotting Distributions](http://www.cookbook-r.com/Graphs/Plotting_distributions_%28ggplot2%29/)

*Lecture Notes*

Chapter 3, Plotting Distributions and Condtional Means: `03-plot_means.Rmd`.

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Final Project*
First progress reports due, Sunday midnight

*Synchronous session: Presenting results in graphical format: barplots, density plots, dot plots, histograms*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i` Topic for the Week: Getting Data --pre-processed data, flat files, basic concepts of "tidy data"

*Resources*

Wickham: [Data import](http://r4ds.had.co.nz/data-import.html), [Tidy data](http://r4ds.had.co.nz/tidy-data.html)

*Lecture Notes*

Chapter 4, Flat Data `04-flat_data.Rmd`

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Synchronous session: working with various data formats*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i` Topic for the Week: Analyzing Data--Linear Regression

*Resources*

Wickham: [Model: Introduction](http://r4ds.had.co.nz/model-intro.html),  [Model Basics](http://r4ds.had.co.nz/model-basics.html), [Model Building](http://r4ds.had.co.nz/model-building.html)

*Lecture Notes*

Chapter 5, Linear Regression `05-regression.Rmd`

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Synchronous session: using linear regression, training and testing Models*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i` Topic for the Week: Presenting data--scatterplots

*Resources*

Wickham: [Data Visualization](http://r4ds.had.co.nz/data-visualisation.html), [Graphics for Communication](http://r4ds.had.co.nz/graphics-for-communication.html)

Tufte, Visual Display chapters 4 and 5. 

Tufte, Envisioning Information, chapter 2

*Lecture Notes*

Chapter 6, Scatterplots `06-scatterplots.Rmd`

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Final Project*

Second Progress Report for Final Project Due

*Synchronous session: Presenting Data via Scatterplots*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i` Topic for the week: Getting Data-- Getting data from the web

*Resources*

Rvest Vignette: [https://cran.r-project.org/web/packages/rvest/vignettes/selectorgadget.html](https://cran.r-project.org/web/packages/rvest/vignettes/selectorgadget.html)

[Reed College rvest introduction](http://www.reed.edu/data-at-reed/resources/R/rvest.html)

[rvest tutorial](https://stat4701.github.io/edav/2015/04/02/rvest_tutorial/)

*Lecture Notes*

Chapter 7, Web Scraping and APIs, `07-webscrape.Rmd`

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Synchronous session: Accessing data from the web*

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```
## Week `r i` Topic for the Week: Analyzing Data--Classification

*Resources*

James, G., Witten, D., Hastie, T., & Tibshirani, R. (2013). An introduction to statistical learning (Vol. 6). New York: Springer. [Chapter 4](http://www-bcf.usc.edu/~gareth/ISL/ISLR%20Sixth%20Printing.pdf) , 
[Chapter 4 Lab R Code](http://www-bcf.usc.edu/~gareth/ISL/Chapter%204%20Lab.txt)

Althoff, T., Danescu-Niculescu-Mizil, C., & Jurafsky, D. (2014, May). How to Ask for a Favor: A Case Study on the Success of Altruistic Requests. In ICWSM. (Available Here)[http://www.aaai.org/ocs/index.php/ICWSM/ICWSM14/paper/download/8106/8101]

*Lecture Notes*

Chapter 8, Classification, `08-classification.Rmd`

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Synchronous Session: Classification*

Classifying behavior via text analysis: random acts of pizza.

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i` Topic for the Week: Presenting Data-- Plots and Tables for Classification

*Resources*


*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Final Project*

Third Progress Report Due

*Lecture Notes*

Chapter 9, plots and tables for classification `09-plotting_classification.Rmd`

*Synchronous Session: Presenting Data, Plots and Tables for classification*


```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i` Topic for the Week: Analyzing Data-- Cross Validation

*Resources*

Wickham [Many Models](http://r4ds.had.co.nz/many-models.html)

Lecture Notes


```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i` Topic for the Week: Getting Data--databases and relational data

*Resources*
Wickham [Relational Data](http://r4ds.had.co.nz/relational-data.html)

Working with Databases in R, available: [https://cran.r-project.org/web/packages/dplyr/vignettes/databases.html](https://cran.r-project.org/web/packages/dplyr/vignettes/databases.html)

*Lecture Notes*

Chapter 11, Databases `11-databases.Rmd`

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Synchronous session* Databases and relational data

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i` Topic for the Week: Unsupervised learning 

Subtopics: k-means, nearest neighbor clustering

*Resources*

James, G., Witten, D., Hastie, T., & Tibshirani, R. (2013). An introduction to statistical learning (Vol. 6). New York: Springer. [Chapter 10](http://www-bcf.usc.edu/~gareth/ISL/ISLR%20Sixth%20Printing.pdf) , 
[Chapter 10 Lab R Code](http://www-bcf.usc.edu/~gareth/ISL/Chapter%2010%20Labs.txt)

*Lecture Notes*

Chapter 12, Unsupervised Learning `12-unsupervised.Rmd`

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Final Project*

Fourth Progress Report Due


*Synchronous session:* K-means clustering, nearest neighbor classification

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i`: Topic for the Week: Presenting Data-- Interactive Graphics

*Resources*

*Lecture Notes*

Chapter 13, Interactive Graphics `13-interactive.Rmd`

*Problem sets*

Problem set `r i-1` Due Midnight Sunday

*Synchronous Session* Interactive Graphics

```{r echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
i<-i+1
```

## Week `r i`: Topic for the Week: Communicating Results

Problem set `r i-1` Due Midnight Sunday

*Synchronous Sessions* Class Presentations 

## Final Products Due Two Weeks After Last Class